Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-02-11 doc/paper.qmd
docs: :memo: revised introduction based on feedback from Mario

The time when a single scientist could work alone in a lab and make
meaningful progress in their field is long gone. Scientific research in
the modern era, especially in certain fields, almost requires working
with other people, oftentimes a large number of people. Research
questions are increasingly more challenging to answer as they require
more time, resources, expertise, and personnel in order to meaningfully
answer.

With the current needs and limitations such as resources when doing
scientific research, collaborations range from those within a single
group or lab, to a few researchers across different departments in the
same institution, to many researchers and personnel spread across
multiple research centers. Collaborations are needed to address basic
limitations of working with small or constrained populations, to sharing
valuable but expensive equipment, or to assisting with technical
expertise or domain knowledge.

At the same time, there is also a higher need for science to be more
open, higher quality, and more rigorous[@Cole2024, @Umbach2024], which
necessitates that researchers collaborate with more researchers and with
a more diverse group of researchers to fulfill these needs. Open
scientific practices, such as open data, open source, and open
materials, aim to improve the reliability and accessibility of
scientific output to increase the societal impact of the results.

There are many hypothetical benefits to openness in science, including
in collaboration. For instance:

@@

From our own personal experience working in the field of health
research, we have varied and diverse ways of collaborating. Often the
most commonly used approach to collaboration is emailing Word documents
around to get feedback and using meetings to discuss and agree on
things. Rarely do we systematically examine whether we are following
best practices for collaboration nor do we collaborate in an open way
following the principles of open science. This lack of a clear standard
or approach to collaborating effectively and openly can greatly impact
productivity and quality of scientific research. It emphasizes a greater
need for a paradigm shift in how we researchers collaborate to be able
to effectively and efficiently tackle the challenges of the modern era.

For this scoping review, we define open collaboration using the
definition as found in [@Forte2013]:

@@

Open collaboration has a lot of potential benefits, many of which are
the same benefits as seen with open science. For instance, open
collaboration can lead to:

Unfortunately, there are, to our knowledge, very few resources and
examples of how to integrate open science into collaborations and how to
make collaboration be open.

This scoping review will focus on current practices of open
collaboration and open science in relation to collaboration in the field
of biomedical and health research.

>>>>

The time when a single scientist could work alone in a lab and make
meaningful progress in their field is long gone. Scientific research in
the modern era, especially in certain fields, requires working with
other people, oftentimes a large number of people [@Forscher2022].
Research questions are increasingly more challenging to answer as they
require more time, resources, expertise, and personnel in order to
meaningfully answer them.

With the current needs and limitations when doing scientific research,
collaborations range from those within a single group or lab, to a few
researchers across different departments in the same institution, to
many researchers and personnel spread across multiple research centers.
Collaborations are needed to address basic limitations of working with
small or constrained populations, to sharing valuable but expensive
equipment, or to assisting with technical expertise or domain knowledge.

At the same time, there is also a higher need for science to be more
open, higher quality, and more rigorous [@Cole2024; @Umbach2024;
@Korbmacher2023], which necessitates that researchers collaborate with
more researchers and with a more diverse group of researchers to fulfill
these needs. Open scientific practices, such as open data, open source,
and open materials, aim to improve the reliability and accessibility of
scientific output to increase the societal impact of the results.
There's greater attention and scrutiny towards verification of results,
a core tenet of science, through replication and reproducibility. In
many ways, the "replication crisis" in science has been a driving force
behind the push for more open science and "team science" practices that
necessitate more collaboration [@Korbmacher2023].

There are many claimed benefits to openness in science, including in
collaboration [@Forscher2022; @Kavanagh2021]. For instance:

@@

From our own personal experience working in the field of health
research, we have varied and diverse ways of collaborating and of being
open. Often the most commonly used approach to collaboration is emailing
Word documents around to get feedback and using meetings to discuss and
agree on things. Rarely do we systematically examine whether we are
following best practices for collaboration nor do we collaborate in an
open way following the principles of open science. As for open science,
while it was first introduced in 1985 [@Chubin1985] and discussed in
academic outputs regularly since then, the actual practices of being
open in science have been minimal or largely confined to specific
sub-branches of openness like open access.

This lack of visible role-models or standards of practice for
effectively collaborating and actively working openly presents a clear
barrier to uptake of these practices. This has real consequences to
continued productivity and quality of scientific research, which is
increasingly apparent in the modern era. These consequences are even
more apparent when comparing to more technically-inclined communities
such as those in open source software. These communities have massively
benefitted from integrating more openness into their workflows, to the
extent that the vast majority of the world's digital infrastructure is
built on and runs open source software. This highlights a greater need
for a paradigm shift in how we researchers collaborate to be able to
effectively and efficiently tackle the challenges of the modern era.

Open collaboration has many potential benefits, of which most overlap
with the benefits of open science. For instance, open collaboration can
lead to:

Unfortunately, there are, to our knowledge, very few resources and
examples of how to integrate open science into collaborations and how to
make collaboration be open [@Korbmacher2023].

Our aim is to conduct a scoping review to provide an overview of current
practices of open collaboration in the academic world. For this scoping
review, we define open collaboration using the definition as found in
[@Forte2013]:

@@
